goal is to understand what types of selections gini and l1 are making

# simulated

```{r}
emlg <- read_rds(here("data-documentation/emlg.rds"))
```

the thing to notice is that 

- gini takes the median of the dominant group (not totally clear how it chooses)

- l1 anywhere in the median (here its even so its a range)

```{r}
emlg %>% emlg_to_gini_l1()

sort(emlg$difference_in_easy)
```

# create

```{r}
emlg <- 
    tibble(
        ref_easy = c(0, 0, 0, 0, 0.5, 1, 1, 1, 1),
        foc_easy = rep(0, 9)
    ) %>% 
    mutate(difference_in_easy = ref_easy - foc_easy)
```

- gini again the median of the dominant group

- l1 chooses the median

```{r}
emlg %>% emlg_to_gini_l1()

sort(emlg$difference_in_easy)
```

# create

```{r}
emlg <- 
    tibble(
        ref_easy = seq(0, 1, length.out = 9),
        foc_easy = rep(0, 9)
    ) %>% 
    mutate(difference_in_easy = ref_easy - foc_easy)
```

- gini again the median of the dominant group but its not clear how it finds it here but it does 

- l1 chooses the median as always

```{r}
emlg %>% emlg_to_gini_l1()

sort(emlg$difference_in_easy)
```

